---
title: "classify-raster-data"
author: "Kristin Braziunas"
date: "June 21, 2016"
output: html_document
---

## Learning how to classify raster data

```{r import-libraries }

# load libraries
library(raster)
library(rgdal)

```

## Import CHM (Canopy Height Model)

```{r import-canopy-height }

# import TEAK CHM
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

# test plot
plot(chm, main = "Plot TEAK CHM using raster package")

# test stretch image
image(log(chm), main = "Plot TEAK CHM image \nJust render pixels that stretch to fill the space")

```

## Deal with 0 values

```{r assign-values-na }

# plot hist to identify 0 values
hist(chm)

# set 0 values to NA
chm[chm == 0] <- NA

# look at hist again
hist(chm, xlab = "Tree height (m)")  # lots of tall trees

```

## Read in aspect layer

```{r import-aspect-data }

# read in aspect
aspect <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarAspect.tif")

# plot aspect
plot(aspect, main = "Plot of TEAK aspect data")

```

## Let's look at how aspect relates to tree height

Are trees taller on north or south aspects?

First, we will set north-facing slopes to 1 and south-facing slopes to 2.
* North-facing slopes are 0-45 and 315-360 degrees, Class = 1
* South-facing slopes are 135-225 degrees, Class = 2

## Create classification matrix

```{r class-matrix }

# create array of values
class.m <- c(0, 45, 1,
             45, 135, NA,
             135, 225, 2,
             225, 315, NA,
             315, 360, 1)

# shape values into matrix
rcl.m <- matrix(class.m,
                ncol = 3,
                byrow = TRUE)

# let's look at our matrix
rcl.m

```

## Now reclassify the raster

```{r perform-reclass }

# reclassify raster using matrix
asp.ns <- reclassify(aspect,
                     rcl.m)

# plot reclassified raster
plot(asp.ns, main = "North- and south-facing slopes")

```

## Export Geotiff

Save reclassified matrix as output.

```{r export-tiff, eval=FALSE }

writeRaster(asp.ns,
            file = "../outputs/TEAK/TEAK_nsAspect2.tif",
            options = "COMPRESS=LZW", 
            NAflag = -9999)

# compress to keep file size as small as possible

```

## Create a mask

We want to create a raster mask that will compare two rasters only at the selected points of north- and south-facing slopes. This masks out the data we are not interested in.

Important: When using a mask to compare two different rasters, need to pay attention to:
*Are the two rasters the same resolution?
*Are the locations matched up or is there georeference data in the raster?

```{r create-raster-mask }

# create a mask from asp.ns
asp.ns

# import NDVI layer
ndvi <- raster("../NEONdata/D17-California/TEAK/2013/spectrometer/veg_index/TEAK_NDVI.tif")

# plot it
plot(ndvi,
     main = "NDVI for TEAK site")

# mask data
nsFacing.ndvi <- mask(ndvi, asp.ns)

# plot it
plot(nsFacing.ndvi, 
     main = "NDVI for north- and south-facing slopes \nTEAK field site")

```

## Subset aspect raster to only select north-facing slopes

```{r subset-raster-north }

# subset raster so only north-facing slopes have a value
asp.n <- asp.ns == 1
asp.n[asp.n == 0] <- NA

# test plot
plot(asp.n,
     main = "North-facing slopes \nTEAK field site")

```

## Use this new north-facing raster to mask NDVI

```{r mask-raster-north }

# mask NDVI with north-facing slopes
nFacing.ndvi <- mask(ndvi, asp.n)

# plot
plot(nFacing.ndvi,
     main = "NDVI for north-facing slopes \nTEAK field site")

```

## Subset aspect raster to only select south-facing slopes

```{r subset-raster-south }

# subset raster so only south-facing slopes have a value
asp.s <- asp.ns == 2
asp.s[asp.s == 0] <- NA

# test plot
plot(asp.s,
     main = "South-facing slopes \nTEAK field site")

```

## Use this new south-facing raster to mask NDVI

```{r mask-raster-south }

# mask NDVI with south-facing slopes
sFacing.ndvi <- mask(ndvi, asp.s)

# plot
plot(sFacing.ndvi,
     main = "NDVI for south-facing slopes \nTEAK field site")

```